Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-08-03 documentation/paper.tex
Started documenting iterative cluster nucleation
Added way more details on ICN

\begin{enumerate}
    \item A detailed simulation of the physical environment is often used to produce an accurate discriminator. This takes a long time even for a team of experienced physicists, and will have to be updated whenever the physical variables of the experimental apparatus change.
    \item The primary intention of many dark matter experiments, which is to reduce background radiation to the lowest level possible, means that there will be a very small number of background events detected. This places a severe limit on the amount of data that is available to optimize such a discriminator.
    \item Furthermore, what little data is available to optimize a discriminator very often contains impurities, because whatever background radiation is present during WIMP detection runs is also present during calibration runs. These are difficult to separate without already having access to a functional discriminator, creating a ``chicken or the egg'' problem (from which the most common way to escape is to create a time- and labor-intensive simulation).
\end{enumerate}

The objective of this study is to solve all 3 of these problems using machine learning. This is done using semi-supervised algorithms that simultaneously learn an effective discriminator and improves the purity of its own training data set.

@@

>>>>

\begin{enumerate}
    \item A detailed simulation of the physical environment is often used to produce an accurate discriminator. This takes a long time even for a team of experienced physicists, and will have to be updated whenever the physical variables of the experimental apparatus change.
    \item The primary intention of many dark matter experiments, which is to reduce background radiation to the lowest level possible, means that there will be a very small number of background events detected. This places a severe limit on the amount of data that is available to optimize such a discriminator.
    \item Furthermore, what little data is available to optimize a discriminator very often contains impurities, because whatever background radiation is present during WIMP detection runs is also present during calibration runs. These are difficult to separate without already having access to a functional discriminator, creating a ``chicken or the egg'' problem (from which the most common way to escape is to create a time- and labor-intensive simulation).
    \item When a discriminator is optimized on the impure training data available, for which ground truths are defined based on the calibration run of origin, it is really being optimized to discriminate between calibration runs, rather than between classes of particles. Without careful consideration, these two goals may not match up. For instance, a discriminator can undesirably fit on noise that varies between calibration runs.
\end{enumerate}

The objective of this study is to solve all 3 of these problems using machine learning. This is done using semi-supervised algorithms that simultaneously learn an effective discriminator and improves the purity of their own training data sets.

@@

\section{Semi-Supervised Learning}

\subsection{Iterative Cluster Nucleation}

Iterative cluster nucleation is a semi-supervised learning algorithm that takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator. The algorithm is initialized as follows:

\begin{enumerate}
    \item Take a subset of the training data, referred to as the seed set $\varsigma$. Use this as the beginning of the training set.
    \item Remove classifications from any other available training examples. These create the unlabeled set $\upsilon$.
    \item Compile and randomly initialize the weights of a neural network discriminator $NN(x)$ (where $x$ is the input data format).
\end{enumerate}

Parameterized by the initial seed threshold $j$ (on the order of 0.01) and the seed threshold multiplier $k$ (on the order of 1.05), the algorithm follows these iterative steps:

\begin{enumerate}
    \item Train $NN(x)$ for 30 epochs on $\varsigma$, using the imperfect ground truth values available.
    \item Using the partially trained weights of $NN(x)$, run inference on the entirety of $\upsilon$, producing a set of predictions $p$.
    \item Find predictions within $p$ that are within a distance of $j$ of either 0 or 1; as this is a binary classifier, such a prediction represents high confidence. Remove any such examples from $\upsilon$ and add them to $\varsigma$. The principle is that, when $NN(x)$ is trained on a relatively small set for short period of time, the few predictions within $p$ that are very confident are highly likely to be correct.
    \item If no examples have been removed from $\upsilon$ and added to $\varsigma$, multiply $j$ by $k$ in place. This is done to increase the acceptance rate later in the training process, when most easily classifiable examples have been added to $\varsigma$. Otherwise, gridlock would occur, where certain examples could not be confidently classified given the training set.
\end{enumerate}

To arrive at an effective discriminator, the network first trains on a smaller amount of imperfect data. Because the ground truth data is imperfect, training on a relatively small set for a short period of time (potentially using regularization, in addition) ensures that the network will be less likely to overfit and its learned function will be simpler. This is in line with the objective, since, assuming the training inputs are normalized correctly, it should be simpler to discriminate between classes of particles than calibration runs.

The essence of this algorithm, then, is a positive feedback loop in which, by including the network's most confident predictions in the training set, the simpler function that is desired is reinforced, even when the original ground truths teach the more complex function that would likely have been arrived at by a conventional supervised learning system.
